### 2022-11-25 paper3.tex
chapter specific referencing

SSD for replication studies comes with unique opportunities and challenges; the
data from the original study can be used to inform SSD, at the same time the
analysis of replication success based on original and replication study is
typically different from an analysis of a single study for which traditional SSD
methodology was developed. Since analysis and design of replication studies
should be in accordance, a relatively small literature has emerged which
specifically deals with replication study power calculations and SSD
\citep{Bayarri2002, Goodman1992, Senn2002, Anderson2017, Micheloud2020,
  vanZwet2022, Held2020, Pawel2022b, Hedges2021, Anderson2022}. However, most of
these articles only deal with selected analysis methods and data models. % For
% instance, SSD for standardized mean difference effect sizes analyzed with Bayes
% factors \citep{Bayarri2002}, SSD for statistical significance assessment of the
% replication \citep{Goodman1992, Senn2002, Micheloud2020, vanZwet2022}, SSD for
% reverse-Bayes assessment of replication studies \citep{Held2020, Pawel2022b}, or
% SSD for meta-analysis of replication studies \citep{Hedges2021}.
An exception is the excellent article by \citet{Anderson2022} which discusses
more general principles of replication SSD in the context of psychological
research, mostly from a frequentist perspective. As they state ``the literature
on Bayesian sample size planning is still nascent, particularly with respect to
Bayes Factors \citep{Schoenbrodt2017}, and has not yet been clearly optimized
for the context of most replication goals'' \citep[p. 18]{Anderson2022}. Our
goal is therefore to complement their article by developing a unified framework
of replication SSD (schematically illustrated in Figure~\ref{fig:SSDschema})
based on principles from Bayesian design approaches \citep{Spiegelhalter1986b,
  Spiegelhalter1986c, Weiss1997, OHagan2001b, Gelfand2002, DeSantis2004,
  Spiegelhalter2004, Schoenbrodt2017, Pek2019, Kunzmann2021, Park2022,
  Grieve2022}. We aim to provide both a theoretical basis for methodologists
developing new methods for design and analysis methods of replication studies,
and also to illustrate how Bayesian design approaches can practically be used by
researchers planning a replication study.

@@

  \caption{Schematic illustration of Bayesian sample size determination for
    replication studies. The original and replication data are denoted by
    $x_{o}$ and $x_{r}$, respectively. Both are assumed to come from a
    distribution with density/probability mass function denoted by
    $f(x_{i} \given \theta)$ for $i \in \{o, r\}$. An initial prior with density
    function $f(\theta \given \text{external knowledge})$ is assigned to the
    model parameter $\theta$.}
\label{fig:SSDschema}
\end{figure}

@@

\section{General framework} \label{sec:GFW}

% Figure~\ref{fig:SSDschema} schematically illustrates the process of Bayesian SSD
% for replication studies, which we will explain in more detail in the following.
Suppose an original study has been conducted and resulted in a data set $x_{o}$.
These data are assumed to come from a distribution characterized by an unknown
parameter $\theta$ and with density function $f(x_{o} \given \theta)$. To assess
the replicability of a claim from the original study, an independent and
identically designed (apart from the sample size) replication study is
conducted, and the goal of the design stage is to determine its sample size
$n_{r}$.

As the observed original data $x_{o}$, the yet unobserved replication data
$X_{r}$ are assumed to come from a distribution depending on the parameter
$\theta$. The parameter $\theta$ thus provides a link between the two studies,
and the knowledge obtained from the original study can be used to make
predictions about the replication. The central quantity for doing so is the
so-called \emph{design prior} of the parameter $\theta$, which we write as the
posterior distribution of $\theta$ based on the original data and an
\emph{initial prior} for $\theta$
\begin{align}
  \label{eq:dp}
  f(\theta \given x_o, \text{external knowledge}) =
  \frac{f(x_o \given \theta) \, f(\theta \given \text{external knowledge})}{f(x_o \given
  \text{external knowledge})}.
\end{align}
The initial prior of $\theta$ may depend on external knowledge (\eg{} data from
other studies) and it represents the uncertainty about $\theta$ before observing
the original data. We will discuss common types of external knowledge in the
replication setting in the next Section. %~\ref{sec:SS}.
The design prior~\eqref{eq:dp} hence represents the state of knowledge and
uncertainty about the parameter $\theta$ before the replication is conducted,
and, along with an assumed replication sample size $n_{r}$, it can be used to
compute a predictive distribution for the replication data
\begin{align}
  \label{eq:ftr}
  f(x_r \given n_r, x_o, \text{external knowledge})
  &= \int%_\Theta
  % \underbrace{f(x_r \given n_{r}, \theta)}_{\text{likelihood}}
    f(x_r \given n_{r}, \theta)
    % \, \underbrace{f(\theta \given x_o, \text{external knowledge})}_{\text{design prior}}
    \, f(\theta \given x_o, \text{external knowledge})
    \, \text{d}\theta.
\end{align}

After completion of the replication, the observed data $x_r$ will be analyzed in
some way to quantify to what extent the original result could be replicated. The
analysis may involve the original data (for example, a meta-analysis of the two
data sets) or it may only use the replication data. Typically, there is a
\emph{success region} $S$ which implies that if the replication data fall within
it ($x_r \in S$), the replication is successful. The \emph{probability of
  replication success} can thus be computed by integrating the predictive
density~\eqref{eq:ftr} over $S$. To ensure a sufficiently conclusive replication
design, the sample size $n_{r}$ is determined such that the probability of
replication success is at least as large as a desired target probability of
success, here and henceforth denoted by $1 - \beta$. The required sample size
$n_r^*$ is then the smallest sample size which leads to a probability of
replication success of at least $1 - \beta$, \ie{}
\begin{align}
  \label{eq:nr}
  n_r^* = \inf \left\{n_r : \Pr(X_{r} \in S \given
  n_r, x_{o}, \text{external knowledge}) \geq 1 - \beta \right\}.
\end{align}

@@

\section{Sample size determination in the normal-normal hierarchical model}
\label{sec:SS}
We will now illustrate the general methodology from the previous section in the
\emph{normal-normal hierarchical model} where predictive distributions and the
probability of replication success can often be expressed in closed-form,
permitting further insight. % To conduct SSD for replication studies
It is pragmatic to adopt a meta-analytic perspective and use only study level
summary statistics instead of the raw study data since the raw data from the
original study are not always available to the replicators. Typically, the
underlying parameter $\theta$ is a univariate effect size quantifying the effect
on the outcome variable (\eg{} a mean difference, a log odds ratio, or a log
hazard ratio). The original and replication study can then be summarized through
an effect estimate $\that$, possibly the maximum likelihood estimate, and a
corresponding standard error $\sigma$, \ie{} $x_{o} = \{\that_{o}, \sigma_{o}\}$
and $x_{r} = \{\that_{r}, \sigma_{r}\}$. Effect estimates and standard errors
are routinely reported in research articles or can, under some assumptions, be
computed from $p$-values and confidence intervals. As in the conventional
meta-analytic framework \citep{Sutton2001}, we further assume that for study
$k \in \{o, r\}$ the (suitably transformed) effect estimate $\hat{\theta}_k$ is
approximately normally distributed around a study specific effect size
$\theta_k$ and with (known) variance equal to its squared standard error
$\sigma_k^2$, here and henceforth denoted by
$\hat{\theta}_k \given \theta_k \sim \Nor(\theta_k, \sigma_k^2)$. The standard
error $\sigma_k$ is typically of the form $\sigma_k = \lambda/\sqrt{n_k}$ with
$\lambda^{2}$ some unit variance and $n_{k}$ the sample size. The ratio of the
original to the replication variance is thus the ratio of the replication to the
original sample size
\begin{align*}
  c = \sigma^2_o/\sigma^2_r = n_r/n_o,
\end{align*}
which is often the main focus of SSD as it quantifies how much the replication
sample $n_{r}$ size needs to be changed compared to the original sample size
$n_{o}$. Depending on the effect size type, this framework might require slight
modifications \citep[see \eg{}][ch. 2.4]{Spiegelhalter2004}.

Assuming a normal sampling model for the effect
estimates~\eqref{eq:hat_theta_k}, as described previously, and specifying an
initial hierarchical normal prior for the study specific effect
sizes~\eqref{eq:theta_k} and the effect size~\eqref{eq:theta}, leads then to the
normal-normal hierarchical model
\begin{subequations}
\label{eq:hierarch-model}
\begin{align}
  \hat{\theta}_k \given \mspace{-1mu} \theta_k &\sim \Nor(\theta_k, \sigma_k^2)
  \label{eq:hat_theta_k} \\
  \theta_k \given \theta \,\,  &\sim \Nor(\theta, \tau^2) \label{eq:theta_k} \\
  \theta \,\, &\sim \Nor(\mutheta,
  \sigmatheta^2). \label{eq:theta}
\end{align}
\end{subequations}
By marginalizing over the study specific effects sizes, the
model~\eqref{eq:hierarch-model} can alternatively be expressed as
\begin{subequations}
\label{eq:hierarch-model2}
\begin{align}
  \hat{\theta}_k \given \mspace{-1mu} \theta &\sim \Nor(\theta, \sigma_k^2 + \tau^{2})
  \label{eq:hat_theta_k2} \\
  \theta  &\sim \Nor(\mutheta,
  \sigmatheta^2) \label{eq:theta2}
\end{align}
\end{subequations}
which is often more useful for derivations and computations. In the following we
will explain how the normal-normal hierarchical model can be used
% is similar to the random effects model typically used in meta-analysis, and
% to incorporate the original data and external knowledge into
for SSD of the replication study.

\subsection{Design prior and predictive distribution}
\label{sec:designpredictive}
The observed original data $x_{o} = \{\that_{o}, \sigma_{o}\}$ can be combined
with the initial prior~\eqref{eq:theta2} %by Bayes' theorem~\eqref{eq:dp}
by standard Bayesian theory for normal prior and likelihood \citep[ch.
3.7]{Spiegelhalter2004} to obtain a posterior distribution for the effect size
$\theta$
\begin{align}
  \label{eq:dpnormal}
  \theta \given \hat{\theta}_o, \sigma^2_o
  \sim
  \Nor\left(
  \frac{\hat{\theta}_o}{1 + 1/g} + \frac{\mutheta}{1 + g},
  \frac{\sigma^2_o + \tau^2}{1 + 1/g} \right)
\end{align}
where $g = \sigmatheta^2/(\sigma^2_o + \tau^2)$ is the \emph{relative prior
  variance}. This posterior serves then as the design prior for predicting the
replication data.

It is interesting to contrast the design prior \eqref{eq:dpnormal} to the
``conditional'' design prior \citep{Micheloud2020}, that is, to assume that the
unknown effect size $\theta$ corresponds to the original effect estimate
$\that_{o}$. This is a standard approach in practice, for instance,
\citet{Opensc2015} determined the sample sizes of its 100 replications under
this assumption. In our framework it implies that the normal design
prior~\eqref{eq:dpnormal} becomes a point mass at the original effect estimate
$\that_{o}$, which can either be achieved through overwhelmingly informative
original data ($\sigma^{2}_{o} \downarrow 0$) along with no heterogeneity
($\tau^{2} = 0$), or through an overwhelmingly informative initial prior
($g \downarrow 0$) centered around the original effect estimate
($\mutheta = \that_{o}$). Both cases show that from a Bayesian perspective the
standard approach is unnatural as it either corresponds to making the standard
error $\sigma_{o}$ smaller than it actually was, or to cherry-picking the prior
based on the data.

Based on the design prior~\eqref{eq:dpnormal}, the predictive distribution of
the replication effect estimate $\that_{r}$ can then be computed by assuming a
replication standard error $\sigma_{r}$ and integrating the marginal density of
the replication effect estimate~\eqref{eq:hat_theta_k2} with respect to the
prior density, leading to
\begin{align}
  \label{eq:fthetar}
  \hat{\theta}_r \given \hat{\theta}_o, \sigma^2_o, \sigma^2_r
  \sim
  \Nor\left(\muthatr =
  \frac{\hat{\theta}_o}{1 + 1/g} + \frac{\mutheta}{1 + g}, \sigmathatr^{2} =
  \, \sigma^2_r + \tau^2 + \frac{\sigma^2_o + \tau^2}{1 + 1/g}\right)
\end{align}
which can again be shown using standard Bayesian theory \citep[ch.
3.13.3]{Spiegelhalter2004}.
% For some analysis methods, replication success is more naturally defined via the
% replication $z$-value $z_{r} = \that_{r}/\sigma_{r}$. In these cases, it can be
% more convenient to use the associated predictive distribution of $z_{r}$
% \begin{align}
%   \label{eq:zr}
%   z_r \given z_o, c%,  z_{\scriptscriptstyle{\theta}},h, g
%   \sim
%   \Nor\left\{\muzr =
%   \sqrt{c} \left(\frac{z_o}{1 + 1/g} +
%   \frac{z_{\scriptscriptstyle{\theta}} \sqrt{g  (1 + h)}}{1 + g}\right),
%   \sigmazr^{2} =
%   \, 1 + c \left( h + \frac{1 + h}{1 + 1/g}  \right) \right\}
% \end{align}
% which only depends on relative parameters, \ie{} the variance ratio
% $c = \sigma^{2}_{o}/\sigma^{2}_{r}$, the original $z$-value
% $z_{o} = \that_{o}/\sigma_{o}$, the prior $z$-value
% $z_{\scriptscriptstyle{\theta}} = \mutheta/\sigma_{\scriptscriptstyle{\theta}}$,
% and the relative heterogeneity $h = \tau^{2}/\sigma^{2}_{o}$.
The design prior~\eqref{eq:dpnormal} and the resulting predictive
distribution~\eqref{eq:fthetar}
% and~\eqref{eq:zr}
depend on the parameters of the initial prior ($\tau^ {2}$, $\mutheta$,
$\sigmatheta^{2}$). We will now explain how these parameters can be specified
based on external knowledge.

\subsection{Incorporating external knowledge in the initial prior}
\label{sec:initialPrior}

@@

\subsubsection{Between-study heterogeneity}
\label{sec:heterogeneity}
The expected degree of between-study heterogeneity can be incorporated via the
variance $\tau^2$ in~\eqref{eq:theta_k}. As $\tau^{2}$ decreases, the study
specific effect sizes become more similar, whereas for increasing $\tau^{2}$
they become more unrelated. If the replicators do not expect any heterogeneity
they can thus set $\tau^{2} = 0$ which will lead to the model collapsing to a
fixed effects model.

@@

\subsubsection{Knowledge about the effect size}
Prior knowledge about the effect size $\theta$ can be incorporated via the prior
mean $\mutheta$ and the prior variance $\sigmatheta^2$ in~\eqref{eq:theta}. For
instance, the parameters may be specified based on a meta-analysis of related
studies \citep{McKinney2021} or based on expert elicitation \citep{OHagan2019}.
The resulting design prior will then contain more information than what was
provided by the original data alone, leading to potentially more efficient
designs. If there is no prior knowledge available, a standard approach is to
specify an (improper) flat prior by letting the variance go to infinity
($\sigmatheta^2 \to \infty$). The resulting design prior will then only contain
the information from the original study.

\subsubsection{Exaggerated original results}
\label{sec:shrinkage}
Potentially exaggerated original results can be counteracted by setting
$\mutheta = 0$ which shrinks the design prior towards smaller effect sizes (in
absolute value) than the observed effect estimate $\that_{o}$. For instance,
replicators could believe that the results from the original study are
exaggerated because there is no pre-registered study protocol available. Even
without such beliefs, weakly informative shrinkage priors may also be motivated
from a ``regularization'' point of view as they can correct for statistical
biases \citep{Copas1983, Firth1993} or prevent unreasonable parameter values
from taking over the posterior in settings with uninformative data
\citep{Gelman2009}.

The amount of shrinkage is determined via the prior variance $\sigmatheta^2$. A
diffuse prior ($\sigmatheta^2 \to \infty$) will lead to no shrinkage, while a
highly concentrated prior ($\sigmatheta^2 \downarrow 0$) will completely shrink
the design prior to a point mass on zero. One option for specifying
$\sigmatheta^2$ is to use an estimate from a corpus of related studies. For
instance, \citet{vanZwet2021} used the Cochrane library of systematic reviews to
specify design priors for hypothetical replication studies of RCTs. If no corpus
is available, a pragmatic alternative is to use the empirical Bayes estimate
based on the original data
\begin{align}
  \label{eq:EBestimate}
  \hat{\sigma}^2_{\scriptscriptstyle{\theta}} = \max\{(\that_{o} - \mutheta)^{2} - \tau^{2} - \sigma^{2}_{o}, 0\}.
\end{align}
The estimate~\eqref{eq:EBestimate} will lead to adaptive shrinkage
\citep{Pawel2020} in the sense that shrinkage is large for unconvincing original
studies (those with small effect estimates in absolute value $|\that_{o}|$
and/or large standard errors $\sigma_{o}$), but disappears as the data become
more convincing (through larger effect estimates in absolute value $|\that_{o}|$
and/or smaller standard errors $\sigma_{o}$).

\subsection{Example: Cross-laboratory replication project}
\label{sec:example}
We will now illustrate the construction of design priors based on data from a
recently conducted replication project \citep{Protzko2020}, see
Figure~\ref{fig:data} for a summary of the data. The data were collected in four
laboratories which, over the course of five years, conducted their typical
social-behavioral experiments on topics such as psychology, communication, or
political science. From the experiments conducted in this period, each lab
submitted four original findings to be replicated. For instance, the original
finding from the ``Labels'' experiment was: ``When a researcher uses a label to
describe people who hold a certain opinion, he or she is interpreted as
disagreeing with those attributes when a negative label is used and agreeing
with those attributes when a positive label is used'' \citep[p.
17]{Protzko2020}, which was based on an effect estimate
$\that_{o} = 0.205$ with 95\% confidence interval from
$0.11$ to $0.3$. For each submitted
original finding, four replication studies were then carried out, one by the
same lab (a \emph{self-replication}) and three by the other three labs (three
\emph{external-replications}).

>>>>

SSD for replication studies comes with unique opportunities and challenges; the
data from the original study can be used to inform SSD, at the same time the
analysis of replication success based on original and replication study is
typically different from an analysis of a single study for which traditional SSD
methodology was developed. Since analysis and design of replication studies
should be in accordance, a relatively small literature has emerged which
specifically deals with replication study power calculations and SSD
\citep{Bayarri2002, Goodman1992, Senn2002, Anderson2017, Micheloud2020,
  vanZwet2022, Held2020, Pawel2022b, Hedges2021, Anderson2022}. However, most of
these articles only deal with selected analysis methods and data models. % For
% instance, SSD for standardized mean difference effect sizes analyzed with Bayes
% factors \citep{Bayarri2002}, SSD for statistical significance assessment of the
% replication \citep{Goodman1992, Senn2002, Micheloud2020, vanZwet2022}, SSD for
% reverse-Bayes assessment of replication studies \citep{Held2020, Pawel2022b}, or
% SSD for meta-analysis of replication studies \citep{Hedges2021}.
An exception is the excellent article by \citet{Anderson2022} which discusses
more general principles of replication SSD in the context of psychological
research, mostly from a frequentist perspective. As they state ``the literature
on Bayesian sample size planning is still nascent, particularly with respect to
Bayes Factors \citep{Schoenbrodt2017}, and has not yet been clearly optimized
for the context of most replication goals'' \citep[p. 18]{Anderson2022}. Our
goal is therefore to complement their article by developing a unified framework
of replication SSD (schematically illustrated in Figure~\ref{fig3:SSDschema})
based on principles from Bayesian design approaches \citep{Spiegelhalter1986b,
  Spiegelhalter1986c, Weiss1997, OHagan2001b, Gelfand2002, DeSantis2004,
  Spiegelhalter2004, Schoenbrodt2017, Pek2019, Kunzmann2021, Park2022,
  Grieve2022}. We aim to provide both a theoretical basis for methodologists
developing new methods for design and analysis methods of replication studies,
and also to illustrate how Bayesian design approaches can practically be used by
researchers planning a replication study.

@@

  \caption{Schematic illustration of Bayesian sample size determination for
    replication studies. The original and replication data are denoted by
    $x_{o}$ and $x_{r}$, respectively. Both are assumed to come from a
    distribution with density/probability mass function denoted by
    $f(x_{i} \given \theta)$ for $i \in \{o, r\}$. An initial prior with density
    function $f(\theta \given \text{external knowledge})$ is assigned to the
    model parameter $\theta$.}
\label{fig3:SSDschema}
\end{figure}

@@

\section{General framework} \label{sec3:GFW}

% Figure~\ref{fig3:SSDschema} schematically illustrates the process of Bayesian SSD
% for replication studies, which we will explain in more detail in the following.
Suppose an original study has been conducted and resulted in a data set $x_{o}$.
These data are assumed to come from a distribution characterized by an unknown
parameter $\theta$ and with density function $f(x_{o} \given \theta)$. To assess
the replicability of a claim from the original study, an independent and
identically designed (apart from the sample size) replication study is
conducted, and the goal of the design stage is to determine its sample size
$n_{r}$.

As the observed original data $x_{o}$, the yet unobserved replication data
$X_{r}$ are assumed to come from a distribution depending on the parameter
$\theta$. The parameter $\theta$ thus provides a link between the two studies,
and the knowledge obtained from the original study can be used to make
predictions about the replication. The central quantity for doing so is the
so-called \emph{design prior} of the parameter $\theta$, which we write as the
posterior distribution of $\theta$ based on the original data and an
\emph{initial prior} for $\theta$
\begin{align}
  \label{eq3:dp}
  f(\theta \given x_o, \text{external knowledge}) =
  \frac{f(x_o \given \theta) \, f(\theta \given \text{external knowledge})}{f(x_o \given
  \text{external knowledge})}.
\end{align}
The initial prior of $\theta$ may depend on external knowledge (\eg{} data from
other studies) and it represents the uncertainty about $\theta$ before observing
the original data. We will discuss common types of external knowledge in the
replication setting in the next Section. %~\ref{sec3:SS}.
The design prior~\eqref{eq3:dp} hence represents the state of knowledge and
uncertainty about the parameter $\theta$ before the replication is conducted,
and, along with an assumed replication sample size $n_{r}$, it can be used to
compute a predictive distribution for the replication data
\begin{align}
  \label{eq3:ftr}
  f(x_r \given n_r, x_o, \text{external knowledge})
  &= \int%_\Theta
  % \underbrace{f(x_r \given n_{r}, \theta)}_{\text{likelihood}}
    f(x_r \given n_{r}, \theta)
    % \, \underbrace{f(\theta \given x_o, \text{external knowledge})}_{\text{design prior}}
    \, f(\theta \given x_o, \text{external knowledge})
    \, \text{d}\theta.
\end{align}

After completion of the replication, the observed data $x_r$ will be analyzed in
some way to quantify to what extent the original result could be replicated. The
analysis may involve the original data (for example, a meta-analysis of the two
data sets) or it may only use the replication data. Typically, there is a
\emph{success region} $S$ which implies that if the replication data fall within
it ($x_r \in S$), the replication is successful. The \emph{probability of
  replication success} can thus be computed by integrating the predictive
density~\eqref{eq3:ftr} over $S$. To ensure a sufficiently conclusive replication
design, the sample size $n_{r}$ is determined such that the probability of
replication success is at least as large as a desired target probability of
success, here and henceforth denoted by $1 - \beta$. The required sample size
$n_r^*$ is then the smallest sample size which leads to a probability of
replication success of at least $1 - \beta$, \ie{}
\begin{align}
  \label{eq3:nr}
  n_r^* = \inf \left\{n_r : \Pr(X_{r} \in S \given
  n_r, x_{o}, \text{external knowledge}) \geq 1 - \beta \right\}.
\end{align}

@@

\section{Sample size determination in the normal-normal hierarchical model}
\label{sec3:SS}
We will now illustrate the general methodology from the previous section in the
\emph{normal-normal hierarchical model} where predictive distributions and the
probability of replication success can often be expressed in closed-form,
permitting further insight. % To conduct SSD for replication studies
It is pragmatic to adopt a meta-analytic perspective and use only study level
summary statistics instead of the raw study data since the raw data from the
original study are not always available to the replicators. Typically, the
underlying parameter $\theta$ is a univariate effect size quantifying the effect
on the outcome variable (\eg{} a mean difference, a log odds ratio, or a log
hazard ratio). The original and replication study can then be summarized through
an effect estimate $\that$, possibly the maximum likelihood estimate, and a
corresponding standard error $\sigma$, \ie{} $x_{o} = \{\that_{o}, \sigma_{o}\}$
and $x_{r} = \{\that_{r}, \sigma_{r}\}$. Effect estimates and standard errors
are routinely reported in research articles or can, under some assumptions, be
computed from $p$-values and confidence intervals. As in the conventional
meta-analytic framework \citep{Sutton2001}, we further assume that for study
$k \in \{o, r\}$ the (suitably transformed) effect estimate $\hat{\theta}_k$ is
approximately normally distributed around a study specific effect size
$\theta_k$ and with (known) variance equal to its squared standard error
$\sigma_k^2$, here and henceforth denoted by
$\hat{\theta}_k \given \theta_k \sim \Nor(\theta_k, \sigma_k^2)$. The standard
error $\sigma_k$ is typically of the form $\sigma_k = \lambda/\sqrt{n_k}$ with
$\lambda^{2}$ some unit variance and $n_{k}$ the sample size. The ratio of the
original to the replication variance is thus the ratio of the replication to the
original sample size
\begin{align*}
  c = \sigma^2_o/\sigma^2_r = n_r/n_o,
\end{align*}
which is often the main focus of SSD as it quantifies how much the replication
sample $n_{r}$ size needs to be changed compared to the original sample size
$n_{o}$. Depending on the effect size type, this framework might require slight
modifications \citep[see \eg{}][ch. 2.4]{Spiegelhalter2004}.

Assuming a normal sampling model for the effect
estimates~\eqref{eq3:hat_theta_k}, as described previously, and specifying an
initial hierarchical normal prior for the study specific effect
sizes~\eqref{eq3:theta_k} and the effect size~\eqref{eq3:theta}, leads then to the
normal-normal hierarchical model
\begin{subequations}
\label{eq3:hierarch-model}
\begin{align}
  \hat{\theta}_k \given \mspace{-1mu} \theta_k &\sim \Nor(\theta_k, \sigma_k^2)
  \label{eq3:hat_theta_k} \\
  \theta_k \given \theta \,\,  &\sim \Nor(\theta, \tau^2) \label{eq3:theta_k} \\
  \theta \,\, &\sim \Nor(\mutheta,
  \sigmatheta^2). \label{eq3:theta}
\end{align}
\end{subequations}
By marginalizing over the study specific effects sizes, the
model~\eqref{eq3:hierarch-model} can alternatively be expressed as
\begin{subequations}
\label{eq3:hierarch-model2}
\begin{align}
  \hat{\theta}_k \given \mspace{-1mu} \theta &\sim \Nor(\theta, \sigma_k^2 + \tau^{2})
  \label{eq3:hat_theta_k2} \\
  \theta  &\sim \Nor(\mutheta,
  \sigmatheta^2) \label{eq3:theta2}
\end{align}
\end{subequations}
which is often more useful for derivations and computations. In the following we
will explain how the normal-normal hierarchical model can be used
% is similar to the random effects model typically used in meta-analysis, and
% to incorporate the original data and external knowledge into
for SSD of the replication study.

\subsection{Design prior and predictive distribution}
\label{sec3:designpredictive}
The observed original data $x_{o} = \{\that_{o}, \sigma_{o}\}$ can be combined
with the initial prior~\eqref{eq3:theta2} %by Bayes' theorem~\eqref{eq3:dp}
by standard Bayesian theory for normal prior and likelihood \citep[ch.
3.7]{Spiegelhalter2004} to obtain a posterior distribution for the effect size
$\theta$
\begin{align}
  \label{eq3:dpnormal}
  \theta \given \hat{\theta}_o, \sigma^2_o
  \sim
  \Nor\left(
  \frac{\hat{\theta}_o}{1 + 1/g} + \frac{\mutheta}{1 + g},
  \frac{\sigma^2_o + \tau^2}{1 + 1/g} \right)
\end{align}
where $g = \sigmatheta^2/(\sigma^2_o + \tau^2)$ is the \emph{relative prior
  variance}. This posterior serves then as the design prior for predicting the
replication data.

It is interesting to contrast the design prior \eqref{eq3:dpnormal} to the
``conditional'' design prior \citep{Micheloud2020}, that is, to assume that the
unknown effect size $\theta$ corresponds to the original effect estimate
$\that_{o}$. This is a standard approach in practice, for instance,
\citet{Opensc2015} determined the sample sizes of its 100 replications under
this assumption. In our framework it implies that the normal design
prior~\eqref{eq3:dpnormal} becomes a point mass at the original effect estimate
$\that_{o}$, which can either be achieved through overwhelmingly informative
original data ($\sigma^{2}_{o} \downarrow 0$) along with no heterogeneity
($\tau^{2} = 0$), or through an overwhelmingly informative initial prior
($g \downarrow 0$) centered around the original effect estimate
($\mutheta = \that_{o}$). Both cases show that from a Bayesian perspective the
standard approach is unnatural as it either corresponds to making the standard
error $\sigma_{o}$ smaller than it actually was, or to cherry-picking the prior
based on the data.

Based on the design prior~\eqref{eq3:dpnormal}, the predictive distribution of
the replication effect estimate $\that_{r}$ can then be computed by assuming a
replication standard error $\sigma_{r}$ and integrating the marginal density of
the replication effect estimate~\eqref{eq3:hat_theta_k2} with respect to the
prior density, leading to
\begin{align}
  \label{eq3:fthetar}
  \hat{\theta}_r \given \hat{\theta}_o, \sigma^2_o, \sigma^2_r
  \sim
  \Nor\left(\muthatr =
  \frac{\hat{\theta}_o}{1 + 1/g} + \frac{\mutheta}{1 + g}, \sigmathatr^{2} =
  \, \sigma^2_r + \tau^2 + \frac{\sigma^2_o + \tau^2}{1 + 1/g}\right)
\end{align}
which can again be shown using standard Bayesian theory \citep[ch.
3.13.3]{Spiegelhalter2004}.
% For some analysis methods, replication success is more naturally defined via the
% replication $z$-value $z_{r} = \that_{r}/\sigma_{r}$. In these cases, it can be
% more convenient to use the associated predictive distribution of $z_{r}$
% \begin{align}
%   \label{eq3:zr}
%   z_r \given z_o, c%,  z_{\scriptscriptstyle{\theta}},h, g
%   \sim
%   \Nor\left\{\muzr =
%   \sqrt{c} \left(\frac{z_o}{1 + 1/g} +
%   \frac{z_{\scriptscriptstyle{\theta}} \sqrt{g  (1 + h)}}{1 + g}\right),
%   \sigmazr^{2} =
%   \, 1 + c \left( h + \frac{1 + h}{1 + 1/g}  \right) \right\}
% \end{align}
% which only depends on relative parameters, \ie{} the variance ratio
% $c = \sigma^{2}_{o}/\sigma^{2}_{r}$, the original $z$-value
% $z_{o} = \that_{o}/\sigma_{o}$, the prior $z$-value
% $z_{\scriptscriptstyle{\theta}} = \mutheta/\sigma_{\scriptscriptstyle{\theta}}$,
% and the relative heterogeneity $h = \tau^{2}/\sigma^{2}_{o}$.
The design prior~\eqref{eq3:dpnormal} and the resulting predictive
distribution~\eqref{eq3:fthetar}
% and~\eqref{eq3:zr}
depend on the parameters of the initial prior ($\tau^ {2}$, $\mutheta$,
$\sigmatheta^{2}$). We will now explain how these parameters can be specified
based on external knowledge.

\subsection{Incorporating external knowledge in the initial prior}
\label{sec3:initialPrior}

@@

\subsubsection{Between-study heterogeneity}
\label{sec3:heterogeneity}
The expected degree of between-study heterogeneity can be incorporated via the
variance $\tau^2$ in~\eqref{eq3:theta_k}. As $\tau^{2}$ decreases, the study
specific effect sizes become more similar, whereas for increasing $\tau^{2}$
they become more unrelated. If the replicators do not expect any heterogeneity
they can thus set $\tau^{2} = 0$ which will lead to the model collapsing to a
fixed effects model.

@@

\subsubsection{Knowledge about the effect size}
Prior knowledge about the effect size $\theta$ can be incorporated via the prior
mean $\mutheta$ and the prior variance $\sigmatheta^2$ in~\eqref{eq3:theta}. For
instance, the parameters may be specified based on a meta-analysis of related
studies \citep{McKinney2021} or based on expert elicitation \citep{OHagan2019}.
The resulting design prior will then contain more information than what was
provided by the original data alone, leading to potentially more efficient
designs. If there is no prior knowledge available, a standard approach is to
specify an (improper) flat prior by letting the variance go to infinity
($\sigmatheta^2 \to \infty$). The resulting design prior will then only contain
the information from the original study.

\subsubsection{Exaggerated original results}
\label{sec3:shrinkage}
Potentially exaggerated original results can be counteracted by setting
$\mutheta = 0$ which shrinks the design prior towards smaller effect sizes (in
absolute value) than the observed effect estimate $\that_{o}$. For instance,
replicators could believe that the results from the original study are
exaggerated because there is no pre-registered study protocol available. Even
without such beliefs, weakly informative shrinkage priors may also be motivated
from a ``regularization'' point of view as they can correct for statistical
biases \citep{Copas1983, Firth1993} or prevent unreasonable parameter values
from taking over the posterior in settings with uninformative data
\citep{Gelman2009}.

The amount of shrinkage is determined via the prior variance $\sigmatheta^2$. A
diffuse prior ($\sigmatheta^2 \to \infty$) will lead to no shrinkage, while a
highly concentrated prior ($\sigmatheta^2 \downarrow 0$) will completely shrink
the design prior to a point mass on zero. One option for specifying
$\sigmatheta^2$ is to use an estimate from a corpus of related studies. For
instance, \citet{vanZwet2021} used the Cochrane library of systematic reviews to
specify design priors for hypothetical replication studies of RCTs. If no corpus
is available, a pragmatic alternative is to use the empirical Bayes estimate
based on the original data
\begin{align}
  \label{eq3:EBestimate}
  \hat{\sigma}^2_{\scriptscriptstyle{\theta}} = \max\{(\that_{o} - \mutheta)^{2} - \tau^{2} - \sigma^{2}_{o}, 0\}.
\end{align}
The estimate~\eqref{eq3:EBestimate} will lead to adaptive shrinkage
\citep{Pawel2020} in the sense that shrinkage is large for unconvincing original
studies (those with small effect estimates in absolute value $|\that_{o}|$
and/or large standard errors $\sigma_{o}$), but disappears as the data become
more convincing (through larger effect estimates in absolute value $|\that_{o}|$
and/or smaller standard errors $\sigma_{o}$).

\subsection{Example: Cross-laboratory replication project}
\label{sec3:example}
We will now illustrate the construction of design priors based on data from a
recently conducted replication project \citep{Protzko2020}, see
Figure~\ref{fig3:data} for a summary of the data. The data were collected in four
laboratories which, over the course of five years, conducted their typical
social-behavioral experiments on topics such as psychology, communication, or
political science. From the experiments conducted in this period, each lab
submitted four original findings to be replicated. For instance, the original
finding from the ``Labels'' experiment was: ``When a researcher uses a label to
describe people who hold a certain opinion, he or she is interpreted as
disagreeing with those attributes when a negative label is used and agreeing
with those attributes when a positive label is used'' \citep[p.
17]{Protzko2020}, which was based on an effect estimate
$\that_{o} = 0.205$ with 95\% confidence interval from
$0.11$ to $0.3$. For each submitted
original finding, four replication studies were then carried out, one by the
same lab (a \emph{self-replication}) and three by the other three labs (three
\emph{external-replications}).
